### 2015-04-03 GS-extension.tex
work out FOCs (cheating with symmetry for now)

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,bbm}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}

\setcounter{secnumdepth}{-1}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{assumption}{Assumption}
\newtheorem{fact}{Fact}
\newtheorem{result}{Result}

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}

\begin{document}

\title{\vskip-0.6in \Large Political Uncertainty}
\author{Kristy Buzard}
\date{\vskip-.1in \today}
\maketitle

On Feb. 12, Sebastian and I agreed to focus efforts on finding a base model to facilitate empirical identification. I am pursuing Groseclose $\&$ Snyder (1996), ``Buying Supermajorities,'' APSR
\begin{itemize}
	\item For each legislator $i$, $v(i) = u_i(x) - u_i(s)$, measured in money; this is the reservation price of $i$
		\begin{itemize}
			\item $x$ is an alternative policy proposal; $s$ is the status quo
			\item WLOG, label legislators so that $v(i)$ is a non-increasing function 
			\item Note legislators only have preferences over how they vote, not over which alternative wins
		\end{itemize}
	\item There are two vote buyers; each prefers to minimize total bribes paid while passing his preferred policy, but each would prefer to concede the issue rather than pay more than his WTP
		\begin{itemize}
			\item $A$ prefers $x$; $W_A$ is $A$'s willingness to pay (WTP) for $x$ measured in money
			\item $B$ prefers $s$; $W_B$ is $B$'s WTP for $s$
		\end{itemize}
	\item Bribe offer functions: $a(i)$ and $b(i)$ are $A$ and $B$'s offers to $i$. Legislators take these bribe offers as given and then vote for the alternative that maximizes their payoff
	\item A moves first; $a(i)$ is perfectly observable to $B$ when he moves
	\item Goal: characterize SPNE in pure strategies
		\begin{itemize}
			\item Assume unbribed legislators who are indifferent vote for $s$; all bribed legislators who are indifferent vote for whoever bribed them last
		\end{itemize}
	\item Assume continuum of legislators on $\left[-\frac{1}{2},\frac{1}{2}\right]$
	\item Assume $W_A$ large enough that $x$ wins in equilibrium (no uncertainty case)
	\item $m + \frac{1}{2}$ is fraction of legislators who vote for $x$ as opposed to the status quo, $s$
	\item Results
	\begin{itemize}
		\item Prop 1: three types of equilibria in which $x$ wins; depend on size of $W_B$
		\item Prop 2: $m^*$ (the optimal coalition size) is unique, and provides three cases parameterizing its size in terms of $W_B$, $v(-\frac{1}{2})$ and $v(m^*)$
		\item Prop 3/4: special case where $v(z) = \alpha - \beta z$
	\end{itemize}
\end{itemize}

\vskip.3in
General thoughts on extension to uncertainty
\begin{itemize}
	\item I think, without uncertainty, you would estimate $m^*$ as a function of the parameters of $v$ and WTP
		\begin{itemize}
			\item It's useful that $m^*$ is unique. Not clear it would extend to case of uncertainty, but I think it's likely so I'm going to assume it for now
		\end{itemize}
	\item I'm pretty sure  this predicts that $B$ should never pay anything when there is no uncertainty, but I don't see where they say it explicitly (I should read more carefully to verify)
		\begin{itemize}
			\item Uncertainty should reverse this, right?
			\item What is uncertainty? Make $v(z)$ stochastic is most natural
				\begin{itemize}
					\item I'm going to start with linear parameterization of $v(z)$ and add uncertainty. For now, take same form for $v(z)$
						\[
						  v(z) = \alpha -\beta \cdot z
						\]
						and take $z$ as a random variable distributed normally with mean $z$ and standard deviation $\sigma_z$. This is unfortunate nomenclature, since later we'll probably want to take this to a standard normal ``Z'' 
							\begin{itemize}
								\item This means that the legislators are ordered on the interval according to their ideal points, but there is acknowledged uncertainty surrounding their preferences 
								\item What is the source of this uncertainty? We take it to vary by legislator and also possibly by issue area when we extend the model to account for that
								\item Uncertainty comes from: \textbf{log-rolling and cross-pressuring; length of voting record; electoral incentives that we can't control for}
							\end{itemize}
				\end{itemize}
		\end{itemize}
\end{itemize}

\vskip.3in
Backward induction (legislature moves last; B makes last bribe; A makes first bribe)
\begin{enumerate}
	\item Legislature
		\begin{itemize}
			\item Each legislator $z$ will decide whether to vote for $x$ or $s$ given $z$, $a(z)$ and $b(z)$. Votes for $x$ if
			  \[
				  v(z) = \alpha -\beta \cdot z + a(z) - b(z) > 0
				\]
				(whether the inequality is weak or strict depends on tie-breaking rules set out in the paper)
					\begin{itemize}
						\item Let's start out by thinking of $z$ as being a random variable distributed $N(z,\sigma_z)$
						\item Notice that a bribe from $B$ is subtracted because it makes the alternative \textit{less} attractive
					\end{itemize}
			\item Payoff for vote buyer A if $x$ wins is $U_A(x) - \int_{-\frac{1}{2}}^\frac{1}{2} a(j) dj$
			\item Payoff for vote buyer A if $s$ wins is $U_A(s) - \int_{-\frac{1}{2}}^\frac{1}{2} a(j) dj$
			\item I think the cleanest way to write the condition for whether $x$ wins is
				\[
				  \int_{-\frac{1}{2}}^\frac{1}{2} \mathbbm{1}\left[v(j) \geq 0 \right] dj \geq \frac{1}{2}
				\]
		\end{itemize}
	\item Vote buyer B
		\begin{itemize}
			\item GS assumption on vote buyers' objective is ``each prefers to minimize total bribes paid while passing his preferred policy, but each would prefer to concede the issue rather than pay more than his WTP''
			\item This has to be adapted to our situation with uncertainty
				\begin{itemize}
					\item Groseclose and Snyder formulation would suggest something like
						\[
							\min_{b(z)} \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \hskip.2in \text{subject to} \hskip.2in \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \leq U_B(s) - U_B(x)
						\]
						\[
							\text{and however we write the ``while passing his preferred policy'' constraint}
						\]
					\item With asymmetric $v(z)$ or WTP parameters, it would be easy to get equilibria where only $A$ or $B$ buys votes. But we also have lots of outcomes where both buy votes.
					\item They can easily both have positive probability of winning. But what do we need for this to be an equilibrium in this three stage game?
					\item Uncertainty buys us a lot: no longer this knife edge condition of $A$ pushing to the point that $B$ buys no votes
				\end{itemize}
			\item Let's start with the simple maximize expected value of winning (WTP times probability of winning) net of bribes:
						\begin{multline}
							\max_{b(j)} \ W_B\left\{ \left[ \int_{-\frac{1}{2}}^\frac{1}{2} \mathbbm{1}\left[\alpha -\beta \cdot j + a(j) - b(j) \leq 0 \right] dj \right]\geq \frac{1}{2} \right\} - \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \\
							\text{subject to} \hskip.2in \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \leq W_B \hskip.2in \text{and} \hskip.2in b(j)\geq0 \ \forall j
						\end{multline}
				\begin{itemize}
					\item Problem is that the indicator function will never equal 1 when $J \sim N(j,\sigma_j)$
					\item So what's a reasonable objective function? 
						\begin{itemize}
							\item Maximizing the total probability mass where $v(j) \leq 0$?
							  \begin{multline}
									\max_{b(j)} \ W_B \left[ \int_{-\frac{1}{2}}^\frac{1}{2} \Pr \left[\alpha -\beta \cdot j + a(j) - b(j) \leq 0 \right] dj \right] - \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \\
									\text{subject to} \hskip.2in \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \leq W_B  \hskip.2in \text{and} \hskip.2in b(j)\geq0 \ \forall j
								\end{multline}
									\begin{itemize}
										\item Note the non-negativity constraint means you can't extract money from a legislator who is past the FOC point in order to reallocate to another legislator
										\item Some legislators who are past FOC point will not be lobbied
									\end{itemize}
							\item Suppressing the constraint for now and using the fact that $J \sim N(j,\sigma_j)$, this can be re-written as
								\begin{equation}
									\max_{b(j)} \ W_B \left[ \int_{-\frac{1}{2}}^\frac{1}{2} \Pr \left[\frac{\alpha + a(j) - b(j)}{\beta} \leq j \right] dj \right] - \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj
								\end{equation}
								or
								\begin{equation}
									\max_{b(j)} \ W_B \left[ \int_{-\frac{1}{2}}^\frac{1}{2} \Pr \left[\frac{\frac{\alpha + a(j) - b(j)}{\beta} - j}{\sigma_j} \leq z \right] dj \right] - \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj
								\end{equation}
								where $Z \sim N(0,1)$. \\
							\item	So, this can be written as
								\begin{multline}
									\max_{b(j)} \ \int_{-\frac{1}{2}}^\frac{1}{2} \left[ W_B \left(1- \Phi \left(\frac{\frac{\alpha + a(j) - b(j)}{\beta} - j}{\sigma_j} \right)\right) - b(j) \right] dj \\
									\text{subject to} \hskip.2in \int_{-\frac{1}{2}}^\frac{1}{2} b(j) dj \leq W_B  \hskip.2in \text{and} \hskip.2in b(j)\geq0 \ \forall j
									\label{ex:objB}
								\end{multline}
							\item If the realizations of $Z$ are i.i.d. and the bribes are independent across legislators, (I believe) this can be maximized pointwise--except for the constraint
							\item The FOC for each $j$ is then
								\begin{equation}
								  - W_B \cdot \frac{ \partial \Phi \left(\frac{\frac{\alpha + a(j) - b(j)}{\beta} - j}{\sigma_j} \right)}{\partial b(j)} = 1
									\label{eq:foc}
								\end{equation}
								as long as the constraint is satisfied.
									\begin{itemize}
										\item I don't know how decisions are made if the constraint binds---how the rationing happens
										\item If the constraint is not binding, vote buyer B will pay all legislators (we could put in some kind of fixed cost of dealing with each legislator to short circuit this outcome) except those who are already so favorable to him that they are past the FOC point
										\item Perhaps there's something in equilibrium I'm not seeing yet that will reduce further the number of legislators that are paid
									\end{itemize}
						\end{itemize}
				\end{itemize}
			\item Let's examine the FOC, assuming pointwise maximization is okay as in Equation~\ref{eq:foc}
				\begin{itemize}
					\item It's easier for me to think of the CDF in terms of the non-standardized Normal distribution before taking the derivative. Then, if my math is right, the FOC becomes
						\begin{equation}
							  \frac{1}{\beta}\frac{1}{\sigma_j \sqrt{2\pi}} \: e^{- \frac{\left(\frac{\alpha + a(j) - b(j)}{\beta} - j\right)^2}{2\sigma_j^2} }= \frac{1}{W_B} 
								\label{eq:focB2}
						\end{equation}
							\begin{itemize}
								\item This just says the pdf has to equal $\beta$ divided by the willingness to pay
								\item Rearranging to solve for $b(j)$:
									\begin{equation}
							  		e^{- \frac{\left(\frac{\alpha + a(j) - b(j)}{\beta} - j\right)^2}{2\sigma_j^2} }= \frac{\beta\sigma_j \sqrt{2\pi}}{W_B} 
									\end{equation}
									\begin{equation}
							  		- \left(\frac{\alpha + a(j) - b(j)}{\beta} - j\right)^2 =2\sigma_j^2 \ln\left(\frac{\beta\sigma_j \sqrt{2\pi}}{W_B} \right)
									\end{equation}
								Since the log statement on the RHS will almost certainly be negative, I can take the negative sign to the right and take the square root without worrying about imaginary numbers
								  \begin{equation}
							  		\frac{\alpha + a(j) - b(j)}{\beta} - j = \pm \sqrt{-2\sigma_j^2 \ln\left(\frac{\beta\sigma_j \sqrt{2\pi}}{W_B} \right)}
									\end{equation}
										\begin{itemize}
											\item Notice there are two roots: could be to the left of the mode (less than 0.5 cumulative probability), or to the right (more than 0.5 cumulative probability)
										\end{itemize}
									\begin{equation}
							  		\alpha + a(j) - b(j) = \pm \beta\sigma_j \sqrt{-2 \ln\left(\frac{\beta\sigma_j \sqrt{2\pi}}{W_B} \right)} + j \beta
									\end{equation}
									\begin{equation}
							  		b(j) = \alpha + a(j)  - j \beta \pm \beta\sigma_j \sqrt{-2 \ln\left(\frac{\beta\sigma_j \sqrt{2\pi}}{W_B} \right)}
									\end{equation}
									\begin{equation}
							  		b(j) = \alpha + a(j)  - \beta j \pm \beta\sigma_j \sqrt{2 \left(\ln W_B - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)}
										\label{eq:focB}
									\end{equation}
								\item So, the bribe for legislator $j$ takes legislator j's preferences back to zero in expectation (the first three terms) and then relative to that, either adds or subtracts a term that involves WTP, variance and $\beta$
									\begin{itemize}
										\item Do we have a way of distinguishing between the two? They have very different implications for the CDF, even though the same implications for the pdf
										\item It makes more intuitive sense that the bribe would be increasing in $W_B$ (WTP)
										\item Perhaps we derive the different implications and let the data speak
										\item All the comparative statics will come out with same magnitude, both signs
										\item SOC is obviously okay (essentially $-x e^{\frac{x^2}{2}}$)
									\end{itemize}
							\end{itemize}
					\item How do we deal with this r.v. with continuous support when the range of ideal points is posited to be finite ($[-\frac{1}{2},\frac{1}{2}]$)?

				\end{itemize}
		\end{itemize}	
	\item Vote buyer A
		\begin{itemize}
			\item Whatever we decide for vote buyer B will be the same for vote buyer A
			\item Adapting Expression~\ref{ex:objB} for Vote Buyer A
				\begin{multline}
					\max_{a(j)} \ \int_{-\frac{1}{2}}^\frac{1}{2} \left[ W_A \cdot \Phi \left(\frac{\frac{\alpha + a(j) - b(j)}{\beta} - j}{\sigma_j} \right) - a(j) \right] dj \\
						\text{subject to} \hskip.2in \int_{-\frac{1}{2}}^\frac{1}{2} a(j) dj \leq W_A  \hskip.2in \text{and} \hskip.2in a(j)\geq0 \ \forall j
							\label{ex:objA}
				\end{multline}
					\begin{itemize}
						\item Major problem if we solve this game sequentially: when substituting the FOC for Vote Buyer B (Equation~\ref{eq:focB}) into Vote Buyer A's objective function, the $a(j)$'s cancel out ($b(j)$ is linear in $a(j)$) so the benefit term is constant in $a(j)$
						\item Static best response
						  \begin{equation}
								  W_A \cdot \frac{ \partial \Phi \left(\frac{\frac{\alpha + a(j) - b(j)}{\beta} - j}{\sigma_j} \right)}{\partial a(j)} = 1
									\label{eq:focA}
							\end{equation}
						\item Math looks exactly the same as Eqn. \ref{eq:focB2} to \ref{eq:focB} except with $W_A$ replacing $W_B$ and all the opposite signs, so
							\begin{equation}
								a(j) = -\alpha + b(j)  + \beta j \pm \beta\sigma_j \sqrt{2 \left(\ln W_A - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)}
									\label{eq:focA3}
							\end{equation}		
					\end{itemize}
				\item Plugging either one of the vote buyers' FOCs into the other just implies that $W_A = W_B$, but doesn't say anything about the level of bribes. Use some kind of pareto ranking to choose one?
					\begin{itemize}
						\item Note that to get a solution, I've taken both FOCs to have the same root (either both positive or both negative; Otherwise, we need a very special combination of parameters to get an equilibrium)
						\item Intuition for this equilibrium: each vote buyer exactly brings legislator $j$ back to neutral (compensating both for inherent preference $\alpha - \beta \cdot j$) and the other vote buyer's bribe; they also each tack on a term related to WTP, $\beta$, and $\sigma_j$, but these terms cancel out as well as long as $W_A = W_B$. If $W_A \neq W_B$, then there is not such an interior equilibrium as far as I can tell. 
							\begin{itemize}
								\item Need to look for more equilibria, perhaps in mixed strategies.
								\item Especially a way to get asymmetric pure strategies (I'm happy to have asymmetry between $W_A$ and $W_B$ if necessary)
							\end{itemize}
					\end{itemize}
		\end{itemize}
\end{enumerate}

\newpage
Discrete version (Conversation between Kristy and Frank, 3/27/15)
\begin{itemize}
	\item Purpose: try to understand what the right objective function is. Does maximizing total probability also maximize the probability of winning?
		\begin{itemize}
			\item Note that this model \textit{may} introduce interdependence between the bid functions that does not exist in the continuous model
			\item What is not clear is whether this happens when the WTP constraint is not binding, which I'm going to assume from the start, and maybe impose later
		\end{itemize}
	\item Frank prefers to write
	  \[
		  v(z) = \alpha -\beta z + \ve_z + a(z) - b(z)
		\]
			\begin{itemize}
				\item Note that, in a probit we will have to worry about separately identifying the parameters. So for now, will assume $\ve$ does not vary in $z$
				\item Also going to ignore Vote Buyer A's choice for now, so we have
			\end{itemize}
	  \[
		  v(z) = \alpha -\beta z + \ve - b(z)
		\]
	\item Then the probability that legislator $z$ votes against the new proposal is
	  \[
		  \Pr\left[v(z)\leq 0 \right] = \Pr\left[\alpha -\beta z + \ve - b(z) \leq 0 \right] 
			                            = \Pr\left[\ve \leq \beta z - \alpha + b(z) \right] 
		\]
		Assuming $\ve \sim N(0,1)$, the probability that legislator $z$ votes ``no'' is $\Phi\left(\beta z - \alpha + b(z) \right)$
	\item Take a model with three legislators with ideal points at $z=-\frac{1}{2}$, $z=0$ and $z=\frac{1}{2}$. Then the three probabilities are
		\begin{itemize}
			\item $\Phi\left(-\frac{\beta}{2} - \alpha + b(z) \right)$, $\Phi\left(- \alpha + b(z) \right)$, and $\Phi\left(\frac{\beta}{2} - \alpha + b(z) \right)$
			\item Whether a legislator $z$ votes ``no'' is a random variable, denote it $X(z)$, distributed Bernoulli with this probability:
				\[
				  X(z) \sim \text{Bernoulli}(p(z)) \hskip.2in \text{where} \hskip.2in p(z) = \Phi\left(\beta z - \alpha + b(z) \right)
				\]
			\item Need to max probability of at least two voting ``no'': $\Pr(S \geq 2)$ where $S = X\left(-\frac{1}{2}\right) + X\left(0\right) + X\left(\frac{1}{2}\right)$
			\item Thus, maximization problem for Vote Buyer B (in absence of Vote Buyer A) is
			  \begin{multline}
			    \max_{b\left(-\frac{1}{2}\right), b\left(0\right), b\left(\frac{1}{2}\right)} 
					\Pr\left(X\left(-\frac{1}{2}\right)=1\right)\Pr\left(X\left(0\right)=1\right)\left(X\left(\frac{1}{2}\right)=0\right)  + \\
					\Pr\left(X\left(-\frac{1}{2}\right)=1\right)\Pr\left(X\left(0\right)=0\right)\Pr\left(X\left(\frac{1}{2}\right)=1\right) + \\
					\Pr\left(X\left(-\frac{1}{2}\right)=0\right)\Pr\left(X\left(0\right)=1\right)\Pr\left(X\left(\frac{1}{2}\right)=1\right) + \\
					\Pr\left(X\left(-\frac{1}{2}\right)=1\right)\Pr\left(X\left(0\right)=1\right)\Pr\left(X\left(\frac{1}{2}\right)=1\right)
				\end{multline}
			Substituting from the definition of $X$,
			  \begin{multline}
			    \max_{b\left(-\frac{1}{2}\right), b\left(0\right), b\left(\frac{1}{2}\right)} 
					\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right)\Phi\left(- \alpha + b(0) \right)\left(1-\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right)\right)  + \\
					\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right)\left(1-\Phi\left(- \alpha + b(0) \right)\right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) + \\
					\left(1-\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right)\right)\Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) + \\
					\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right)\Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right)
				\end{multline}
			The FOC wrt to $b\left(-\frac{1}{2}\right)$ is
				\begin{multline}
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(- \alpha + b(0) \right)\left(1-\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right)\right)  + \\
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \left(1-\Phi\left(- \alpha + b(0) \right)\right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) - \\
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) + \\
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) = 0
				\end{multline}
			Simplifying:
			  \begin{multline}
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(- \alpha + b(0) \right)+ \\
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) \\
					- 2\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) = 0
				\end{multline}
			Further:
			  \begin{multline}
					\frac{1}{\sqrt{2\pi}} \: e^{- \frac{\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}\right)^2}{2}} \\ \left[ \Phi\left(- \alpha + b(0) \right)+  \Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) 
					- 2 \Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) \right] = 0
				\end{multline}
			So I think we have three conditions:
			\begin{equation}
					\Phi\left(- \alpha + b(0) \right)+  \Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) 
					- 2 \Phi\left(- \alpha + b(0) \right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) = 0
			\end{equation}
			\begin{equation}
					\Phi\left(- \alpha + b(0) \right)+  \Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right) 
					- 2 \Phi\left(- \alpha + b(0) \right)\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right) = 0
			\end{equation}
			\begin{equation}
					\Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2}) \right)+  \Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) 
					- 2 \Phi\left(-\frac{\beta}{2} - \alpha + b(-\frac{1}{2})\right)\Phi\left(\frac{\beta}{2} - \alpha + b(\frac{1}{2}) \right) = 0
			\end{equation}
		\end{itemize}
\end{itemize}

\newpage
Comparative statics
\begin{itemize}
	\item Vote buyer B (3/25/15) --- note these are not super useful yet --- they're comparative statics of the best response function, but not of the equilibrium
		\begin{itemize}
			\item Those for $j$, $a(j)$ and $\alpha$ don't involve the root term, so don't differ depending on whether we take the plus or minus version
				\[
					\frac{\partial b(j)}{\partial j} = -\beta \hskip.2in \frac{\partial b(j)}{\partial a(j)} = 1 \hskip.2in \frac{\partial b(j)}{\partial \alpha} = 1
				\]
			\item Those for $\beta$ and $\sigma_j$ \textit{do} involve the root term. I will only display the version for the positive root; there is also a version with all the signs reversed
				\[
				  \frac{\partial b(j)}{\partial \beta} = -j + \sigma_j \sqrt{2 \left(\ln W_B - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)} - \frac{\sigma_j}{\sqrt{2 \left(\ln W_B - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)}}
				\]
				\[
				  \frac{\partial b(j)}{\partial \sigma_j} = \beta \sqrt{2 \left(\ln W_B - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)} - \frac{\beta}{\sqrt{2 \left(\ln W_B - \ln \beta\sigma_j - \ln \sqrt{2\pi} \right)}}
				\]
		\end{itemize}
\end{itemize}

\newpage
Notes from 3/16 Skype chat (Kristy and Sebastian)
\begin{itemize}
	\item Seems like FOC for lobby will boil down to each one buying votes until marginal benefit = marginal cost
	\item May want to use some simplified rule / heuristic for lobby's decision: perhaps they reorder the legislators by their $\pm 2$ std. dev. and make some decision based on that ordering who to lobby
	\item Will some kind of rule that looks like RMSE come out of the math? Is it possible to get anything closed-form at all?
\end{itemize}

\vskip.3in
Notes from 3/19 Skype chat (Kristy and Sebastian)
\begin{itemize}
	\item We can use the data we have to horse-race this model against other hypotheses about how lobbyists distribute bribes
		\begin{itemize}
			\item Allocate equally among legislators
			\item Groseclose Snyder with full information: only one side pays
			\item Our model with one dimension
				\begin{itemize}
					\item Could we use the econometric model to benchmark to one where uncertainty disappears? Or, as it does theoretically, does the model have to change completely?
				\end{itemize}
			\item Our model with multiple dimensions
		\end{itemize}
	\item We can think of this as looking for the effect of uncertainty on prices---we'd be pricing uncertainty relative to a model with uncertainty
		\begin{itemize}
			\item ``a metric in dollars of uncertainty''
			\item this is a model of vote buying under uncertainty
		\end{itemize}
	\item Given how different the environment with uncertainty is, Kristy should explore both the sequential model that parrots GS96 and a simultaneous model that is more like GH94 (menu auction)
	\item When mapping to the data, we're going to want to know from the model whether/when total contributions represent WTP.
		\begin{itemize}
			\item It clearly doesn't in the case of certainty. One side pays nothing; the other pays either as much as is necessary to shut down its opponent, or nothing at all the necessary amount exceeds WTP
			\item May be able to show it doesn't matter, that the proportion of total expenditures is a sufficient statistic
		\end{itemize}
	\item Finish writing the model, and then see if we can use the estimates we already have to write a first, very rough draft
\end{itemize}

\end{document}