### 2016-04-13 presentation.tex
remove pe-shocks content from presentation

%\documentclass{beamer} 
\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

\usecolortheme{beaver}
\useinnertheme{default}
\setbeamertemplate{itemize item}[triangle]
\setbeamertemplate{itemize subitem}[triangle]
\setbeamertemplate{itemize subsubitem}[circle]
\setbeamertemplate{enumerate items}[default]
\setbeamertemplate{blocks}[upper=block head,rounded]
\setbeamercolor{item}{fg=black}
\usefonttheme{serif} %should allow ccfonts to take effect

\usepackage{cite}
\usepackage{times, verbatim,xcolor,bm}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	%\usepackage[boldsans]{concmath}
	\usepackage[euler-hat-accent]{eulervm}

\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^{tw}}}
\newcommand{\ga}{\gamma}
\newcommand{\ve}{\varepsilon}
\newcommand{\ta}{\theta}
\newcommand{\de}{\delta}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}

\newenvironment{changemargin}[2]{% 
  \begin{list}{}{% 
    \setlength{\topsep}{0pt}% 
    \setlength{\leftmargin}{#1}% 
    \setlength{\rightmargin}{#2}% 
    \setlength{\listparindent}{\parindent}% 
    \setlength{\itemindent}{\parindent}% 
    \setlength{\parsep}{\parskip}% 
  }% 
  \item[]}{\end{list}}

	\let\Tiny=\tiny

\title[Lobbying and Legislative Uncertainty\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Lobbying and Legislative Uncertainty}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{April 16, 2016}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

\section{Overview}
\subsection{Preview}
\begin{frame}{The Questions}

\pause
\begin{enumerate}[<+->]
\item When 
	\begin{itemize}
		\item E
	\end{itemize}
	\vskip.1in
\item Can
	\begin{itemize}
		\item G
	\end{itemize}
\end{enumerate}

\end{frame}

\begin{frame}{Some Data}

\pause
\begin{enumerate}[<+->]
\item When 
	\begin{itemize}
		\item E
	\end{itemize}
	\vskip.1in
\item Can
	\begin{itemize}
		\item G
	\end{itemize}
\end{enumerate}

\end{frame}

\begin{frame}{Preview}
\frametitle{P}
\pause
With . I:
\pause
\begin{itemize}[<+->]
	\item e
	\item c
		\begin{itemize}[<+->]
			\item b
			\item t
		\end{itemize}
	\item e
\end{itemize}
\end{frame}

\begin{frame}{Preview}
\frametitle{Results}

\pause
\begin{itemize}[<+->]
	\item S
	\item F
	\item D
		\begin{itemize}
			\item P
			\item M
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Literature}
\pause
\begin{itemize}[<+->]
	\item \textbf{Vote Buying in Legislatures}: 
	\item \textbf{Stochastic Voting}: 
	\item \textbf{Uncertainty}: 
\end{itemize}
\end{frame}

\section{Model}
\subsection{Political Structure}
\begin{frame}{Policy and Politics}

\pause
T
\pause
\begin{itemize}
	\item A
	\pause
	\item B
\end{itemize}

\end{frame}

\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item {\bfseries Who}
		\begin{enumerate}[i.]
			%\pause
			\item Governments 
		\end{enumerate}
	%\pause
	\item \textbf{Then}
		\begin{enumerate}[i.]
			%\pause
			\item A
			%\pause
			\item B
		\end{enumerate}
	%\pause
	\item \textbf{What}
	%\pause
		\begin{enumerate}[i.]
			\item If
		\end{enumerate}
\end{enumerate}
\end{frame}

\subsection{The Players}

\begin{frame}{D}
\pause
  Objective function:
\pause
\[
  W = \mathit{CS}_X(\tau) + \ga(s,e) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]
\vskip-.1in
\pause
\begin{itemize}[<+->]
	\item S
		\begin{itemize}
			\item $s$: 
			\item $e$: 
		\end{itemize}
	\item Optimal 
		\begin{itemize}
			\item Ignores 
			\item Takes 
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Political Pressure}
Two potential sources
\pause
\begin{enumerate}[<+->]
	\item Endogenous effort choice of lobby, $e$
		\begin{itemize}[<+->]
			\item Lobby chooses effort to maximize profits, $\pi(\cdot)$, net of lobbying effort, $e$
			\item Call lobby's optimal effort choice $e^L$
						\[
						  e^L = \max_e \pi(\tau(\ga(e))) - e
						\]
		\end{itemize}
\end{enumerate}

\end{frame}

\section{Results}
\subsection{One Vote Buyer}
\begin{frame}{}

\pause
When :

\pause
\begin{itemize}[<+->]
	\item T
	\item I
	\item C
\end{itemize}
\end{frame}

\begin{frame}{When ...}
Now 
\pause
\begin{itemize}[<+->]
	\item Want 
	\item But 
\end{itemize}

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result...}
  Assume ...
\end{beamerboxesrounded}
\end{frame}

\section{Conclusion}

\begin{frame}{Future Work}
\pause
\begin{itemize}[<+->]
	\item A
	\item I
	\item C
\end{itemize}

\end{frame}

\begin{frame}{Conclusion}
Taking into account
\pause
\begin{itemize}[<+->]
		\item provides 
		\item demonstrates 
		\item helps 
\end{itemize}

\end{frame}

\end{document}